---
title: "TSA FINAL PROJECT"
author: "Sean Smyth (s3704987), Daniel Enoka (s3686502) and Charles Galea (s3688570)"
date: "21/05/2019"
output:
  pdf_document: default
  html_document: default
---
# Source: https://sites.google.com/site/arcticseaicegraphs/

# Load libraries
```{r setup, include=FALSE}
library(TSA)
library(tseries)
library(FSAdata)
library(forecast)
library(fUnitRoots)
library(lmtest)
library(dplyr)
library(tidyr)
library(readr)
```

# Load data set
```{r}
#sea_ice_arctic2 <- read_csv("sea_ice_arctic2.csv")
# View(sea_ice_arctic2)

sea_ice <-  read.csv("C:sea_ice_arctic2.csv")  #sea_ice_arctic2.csv

colnames(sea_ice) <- c('Year', 'Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')

sea_ice <- sea_ice %>% gather('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', key = 'Month', value = 'volume') %>% arrange(Year)
```

# Convert to time series
```{r}
sea_ice_ts <- ts(sea_ice$volume, start = 1985, frequency = 12)
#View(sea_ice_ts)
#is.na(sea_ice_ts)
```

# Plot time series data
```{r}
plot(sea_ice_ts, type = "o", xlab = "Year", ylab = "Sea Ice Volume (1000 km3)", main = "Arctic Sea Ice")
```

## Annotated plot time series data to examine seasonality
```{r}
plot(sea_ice_ts, ylab='Sea Ice Volume (1000 km3)', xlab = 'Year', type = "l", main = "Arctic Sea Ice")
points(y=sea_ice_ts,x=time(sea_ice_ts),pch=as.vector(season(sea_ice_ts)))
```


##From the plot of the output we know:

* Seasonality is present
* There is a downward trend (non-stationary)
* There is no obvious change in variance
* It is not possible to determine behaviour due to the presence of seasonality.


# ACF and PACF plots
```{r}
par(mfrow=c(1,2))
acf(sea_ice_ts, lag.max = 60, main = "ACP Plot")
pacf(sea_ice_ts, lag.max = 60, main = "PACF Plot")
```

* Strong correlations at lags 0.5, 1.0, 1.5, and so on, which are slowly decreasing indicate the existence of a seasonal autocorrelation relationship.
* There is also a substantial correlation that needs to be considered.

* We no need to examine the time series, ACF and PACF plots of sea ice volumes after taking the first difference.

# Differencing to de-trend data
```{r}
sea_ice_diff <- diff(sea_ice_ts)
plot(sea_ice_diff, ylab='First Difference of Sea Ice Volume', xlab='Time', main = "Time Series Plot of the
First Differences of Arctic Sea Ice Volumes")
```

* The first difference of the time series plot indicates that the data has now been de-trended.


# ACP and PACF plots
```{r}
par(mfrow=c(1,2))
acf(as.vector(sea_ice_diff),lag.max = 60, main="Sample ACF of First Differences of Sea Ice Volumes")
pacf(as.vector(sea_ice_diff),lag.max = 60, main="Sample PACF of First Differences of Sea Ice Volumes")
```
* Although the data hase been de-trended, the seasonal pattern still remains in the time series.

* Next considered seasonal differencing.

# Seasonal Differencing

```{r}
sea_ice_diff2 <- diff(sea_ice_diff, lag = 12)
plot(sea_ice_diff2,xlab='Time',ylab='First and Seasonal Difference of Sea Ice Volumes',main=
"Time Series Plot of First and Seasonal Differences of Sea Ice Volumes")
```

# ACF and PACF plots
```{r}
par(mfrow=c(1,2))
acf(as.vector(sea_ice_diff2),lag.max=36,ci.type='ma',main="Sample ACF of First
and Seasonal Differences of Sea Ice Volumes")
pacf(as.vector(sea_ice_diff2),lag.max=36,ci.type='ma',main="Sample PACF of First
and Seasonal Differences of Sea Ice Volumes")
```

* After taking the seasonal difference we remove most of the seasonality in the series and very little autocorrelation remains.

* There is 1 significant lag in the first part of the series (prior to the first period - lag 12)

* The model suggested by this plot is D=1, d=1.

* Therefore, try a ARIMA(0,1,1)x(0,1,1)~12~ model.

# Fit of ARIMA(0,1,1)x(0,1,1)~12~ model
```{r}
m1_sea_ice = arima(sea_ice_ts,order=c(0,1,1),seasonal=list(order=c(0,1,1), period=12))
res.m1 = residuals(m1_sea_ice);
plot(res.m1,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```

# ACF and PACF plots for ARIMA(0,1,1)x(0,1,1)~12~ model
```{r}
par(mfrow=c(1,2))
acf(res.m1, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m1, lag.max = 36, main = "The sample PACF of the residuals")
```


























#------- Residuals Approach ----------------

# Initially fit of SARIMA(0,0,0)x(0,1,0)~12~ model
```{r}
m1.sea_ice = arima(sea_ice_ts,order=c(0,0,0),seasonal=list(order=c(0,1,0), period=12))
res.m1 = residuals(m1.sea_ice);
plot(res.m1,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```

## ACF and PACF plots of residuals for SARIMA(0,0,0)x(0,1,0)~12~ model
```{r}
par(mfrow=c(1,2))
acf(res.m1, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m1, lag.max = 36, main = "The sample PACF of the residuals")
```

* There is now no pattern implying the existence of a seasonal trend.

* However, we do observe the existence of an ordinary trend from the ACF and PACF plots of the residuals.

* We will attempt to remove this ordinary trend by fitting a SARIMA(0,1,0)x(0,1,0)~12~ model.

## Fit of SARIMA(0,1,0)x(0,1,0)~12~ model   <--  m2

```{r}
m2.sea_ice = arima(sea_ice_ts,order=c(0,1,0),seasonal=list(order=c(0,1,0), period=12))
res.m2 = residuals(m2.sea_ice);
plot(res.m2,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```


## ACF and PACF plots for fit of SARIMA(0,1,0)x(0,1,0)~12~ model
```{r}
par(mfrow=c(1,2))
acf(res.m2, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m2, lag.max = 36, main = "The sample PACF of the residuals")
```

* No evidence of an ordinary trend remaining in the residuals.

* There is a decreasing pattern in lags 1, 2, 3,.... in the SARMA conponent of the PACF plot. The correation at lag 1 in the ACF plot is significant. This implies the existence of an SMA(1) component.

* Now tried to fit a SARIMA(0,1,0)x(0,1,1)~12~ model to try to remove the remaining seasonal component in the residuals.


# Fit of SARIMA(0,1,0)x(0,1,1)~12~ model.  <-- m3
```{r}
m3.sea_ice = arima(sea_ice_ts,order=c(0,1,0),seasonal=list(order=c(0,1,1), period=12))
res.m3 = residuals(m3.sea_ice);
plot(res.m3,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```

## ACF AND PACF plots for SARIMA(0,1,0)x(0,1,1)~12~ model.
```{r}
par(mfrow=c(1,2))
acf(res.m3, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m3, lag.max = 36, main = "The sample PACF of the residuals")
```

* The autocorrelations, especially the first seasonal lag (lag 1) in the ACF plot, become insignificant or slightly significant after adding the seasonal component.

* The ACF and PACF plots can be used to determine the orders of the ARMA component since there are no highly significant correlations at lags s, 2s, 3s, 

* The ACF plot displays one significant autocorrelation while the PACF plot has two significant autocorrelations. This suggests a ARMA(1,1) and ARMA(2,1) models.

* We will now fit SARIMA(1,1,0)x(0,1,1)~12~, SARIMA(0,1,1)x(0,1,1)~12~ and SARIMA(1,1,1)x(0,1,1)~12~ models.



# Fit of SARIMA(1,1,0)x(0,1,1)~12~ model
```{r}
m4.sea_ice = arima(sea_ice_ts,order=c(1,1,0),seasonal=list(order=c(0,1,1), period=12))
res.m4 = residuals(m4.sea_ice);
plot(res.m4,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```

#ACF and PACF plots for SARIMA(1,1,0)x(0,1,1)~12~ model
```{r}
par(mfrow=c(1,2))
acf(res.m4, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m4, lag.max = 36, main = "The sample PACF of the residuals")
```



# Fit of SARIMA(0,1,1)x(0,1,1)~12~ model
```{r}
m5.sea_ice = arima(sea_ice_ts,order=c(0,1,1),seasonal=list(order=c(0,1,1), period=12))
res.m5 = residuals(m5.sea_ice);
plot(res.m5,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```

#ACF and PACF plots for SARIMA(0,1,1)x(0,1,1)~12~ model
```{r}
par(mfrow=c(1,2))
acf(res.m5, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m5, lag.max = 36, main = "The sample PACF of the residuals")
```




# Fit of SARIMA(1,1,1)x(0,1,1)~12~ model
```{r}
m6.sea_ice = arima(sea_ice_ts,order=c(1,1,1),seasonal=list(order=c(0,1,1), period=12))
res.m6 = residuals(m6.sea_ice);
plot(res.m6,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```

#ACF and PACF plots for SARIMA(1,1,1)x(0,1,1)~12~ model
```{r}
par(mfrow=c(1,2))
acf(res.m6, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m6, lag.max = 36, main = "The sample PACF of the residuals")
```


* The residuals for the SARIMA(1,1,0)x(0,1,1)~12~, SARIMA(0,1,1)x(0,1,1)~12~ and SARIMA(0,1,1)x(0,1,1)~12~ models contain several significant residual autocorrelations.

* Therefore, try SARIMA(0,1,2)x(0,1,1)~12~ and SARIMA(2,1,0)x(0,1,1)~12~


# Fit of SARIMA(0,1,2)x(0,1,1)~12~ model
```{r}
m7.sea_ice = arima(sea_ice_ts,order=c(0,1,2),seasonal=list(order=c(0,1,1), period=12))
res.m7 = residuals(m7.sea_ice);
plot(res.m7,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```

#ACF and PACF plots for SARIMA(0,1,2)x(0,1,1)~12~ model
```{r}
par(mfrow=c(1,2))
acf(res.m7, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m7, lag.max = 36, main = "The sample PACF of the residuals")
```

# Fit of SARIMA(2,1,0)x(0,1,1)~12~ model
```{r}
m8.sea_ice = arima(sea_ice_ts,order=c(2,1,0),seasonal=list(order=c(0,1,1), period=12))
res.m8 = residuals(m8.sea_ice);
plot(res.m8,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```

#ACF and PACF plots for SARIMA(2,1,0)x(0,1,1)~12~ model
```{r}
par(mfrow=c(1,2))
acf(res.m8, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m8, lag.max = 36, main = "The sample PACF of the residuals")
```


* Therefore we can conclude that the orders are: p=0, d=1, q=1, P=0, D=1, Q=1 and s=12 for the SARIMA(p,d,q)x(P,D,Q)~s~ model.


#------------ Model Fitting ------------

# ML estimates for SARIMA(0,1,1)x(0,1,1)~12~ model
```{r}
m1_sea_ice_ts = arima(sea_ice_ts,order=c(0,1,1), seasonal=list(order=c(0,1,1), period=12),method = "ML")
coeftest(m1_sea_ice_ts)
```

# Conditional Least Squares for SARIMA(0,1,1)x(0,1,1)~12~ model
```{r}
m1_sea_ice_CSS = arima(sea_ice_ts,order=c(0,1,1), seasonal=list(order=c(0,1,1), period=12),method ="CSS")
coeftest(m1_sea_ice_CSS)
```

# ML estimates for SARIMA(1,1,1)x(0,1,1)~12~ model
```{r}
m2_sea_ice_ts = arima(sea_ice_ts,order=c(1,1,1), seasonal=list(order=c(0,1,1), period=12),method = "ML")
coeftest(m2_sea_ice_ts)
```

# Conditional Least Squares for SARIMA(1,1,1)x(0,1,1)~12~ model
```{r}
m2_sea_ice_CSS = arima(sea_ice_ts,order=c(1,1,1), seasonal=list(order=c(0,1,1), period=12),method ="CSS")
coeftest(m2_sea_ice_CSS)
```

# ML estimates for SARIMA(2,1,0)x(0,1,1)~12~ model
```{r}
m3_sea_ice_ts = arima(sea_ice_ts,order=c(2,1,0), seasonal=list(order=c(0,1,1), period=12),method ="ML")
coeftest(m3_sea_ice_ts)
```

# Conditional Least Squares for SARIMA(2,1,0)x(0,1,1)~12~ model
```{r}
m3_sea_ice_CSS = arima(sea_ice_ts,order=c(2,1,0), seasonal=list(order=c(0,1,1), period=12),method ="CSS")
coeftest(m3_sea_ice_CSS)
```

# ML estimates for SARIMA(2,1,1)x(0,1,1)~12~ model
```{r}
m4_sea_ice_ts = arima(sea_ice_ts,order=c(2,1,1), seasonal=list(order=c(0,1,1), period=12),method ="ML")
coeftest(m4_sea_ice_ts)
```

# Conditional Least Squares for SARIMA(2,1,1)x(0,1,1)~12~ model
```{r}
m4_sea_ice_CSS = arima(sea_ice_ts,order=c(2,1,1), seasonal=list(order=c(0,1,1), period=12),method ="CSS")
coeftest(m4_sea_ice_CSS)
```


* The ML estimates were determined to be highly significant. The AR component in the SARIMA(1,1,1)x(0,1,1)~12~ model was insignificant.



#------------- Diagnostic Checking -------------

# Time series plot for standardized residuals for SARIMA(0,1,1)x(0,1,1)~12~ model
```{r}
plot(window(rstandard(m3_sea_ice_ts),start=c(1985, 2)),
ylab='Standardized Residuals',type='o',
main="Residuals from the ARIMA(2,1,0)x(0,1,1))_12 Model")
abline(h=0, col = 'red')
```

* Plot suggests no major abnormalities with this model(except around 1994) although there are several outliers around 2010 that may need to be investigated in more detail.

# ACF plot of standardized residuals for SARIMA(2,1,0)x(0,1,1)~12~ model
```{r}
acf(as.vector(window(rstandard(m3_sea_ice_ts),start=c(1985,2))),
lag.max=36,
main="ACF of Residuals from the ARIMA(2,1,0)x(0,1,1)_12 Model")
```

* Besides several slightly significant autocorrelations there is no sign of violation of the independence of residuals.

# Normality
```{r}
hist(window(rstandard(m3_sea_ice_ts),start=c(1985,2)), xlab = 'Standardized Residuals', ylab = 'Frequency', main = 'Residuals of ARIMA(2,1,0)x(0,1,1)_12 Model')
```

```{r}
qqnorm(window(rstandard(m3_sea_ice_ts),start=c(1985,2)),main="Q-Q plot for Residuals: ARIMA(2,1,0)x(0,1,1)_12 Model")
qqline(window(rstandard(m3_sea_ice_ts),start=c(1985,2)))
```

## Fit of SARIMA(0,1,1)x(0,1,1)~12~ model   <--  m1
## Fit of SARIMA(0,1,0)x(0,1,0)~12~ model   <--  m2
## Fit of SARIMA(2,1,0)x(0,1,1)~12~ model   <--  m3

```{r}
par(mfrow=c(1,2))
qqnorm(window(rstandard(m3_sea_ice_ts),start=c(1985,2)),main="ARIMA(2,1,0)x(0,1,1)_12 Model")
qqline(window(rstandard(m3_sea_ice_ts),start=c(1985,2)),v=-2,h=-2)

qqnorm(window(rstandard(m1_sea_ice_ts),start=c(1985,2)),main="ARIMA(2,1,0)x(0,1,1)_12 Model", col="red")
qqline(window(rstandard(m1_sea_ice_ts),start=c(1985,2)),v=2, h=2)
```


```{r}

qqnorm(window(rstandard(m1_sea_ice_ts),start=c(1985,2)),main="Q-Q plot for Residuals: ARIMA(2,1,0)x(0,1,1)_12 Model")
qqline(window(rstandard(m1_sea_ice_ts),start=c(1985,2)))

```



## We'll first fit a plain model with only the first seasonal difference with order D = 1 
## and see if we can get rid of the seasonal trend effect by inspecting the autocorrelation 
## structure of the residuals.

```{r}
ml.sea_ice <- arima(sea_ice_ts, order = c(0,0,0), seasonal = list(order=c(0,1,0), period=12))
res.m1 = residuals(ml.sea_ice);  
par(mfrow=c(1,1))
plot(res.m1,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")
```


```{r}
par(mfrow=c(1,2))
acf(res.m1, lag.max = 36, main = "The sample ACF of the residuals")
pacf(res.m1, lag.max = 36, main = "The sample PACF of the residuals")
```

# From the time series plot, we can conclude that we got rid of the trend. 
# Seasonal autocorrelations are seen clearly in 
# ACF and PACF now at the lags corresponding to the periods. 
# We have one significant correlation at the first seasonal lag in both ACF and PACF. 

##----------------------------

```{r}
par(mfrow=c(1,2))
acf(as.vector(sea_ice_ts),lag.max=36, main = "Sample ACF of Carbon Dioxide Levels")
pacf(as.vector(sea_ice_ts),lag.max=36, main = "Sample ACF of Carbon Dioxide Levels")

```

```{r}

m1_sea_ice = arima(sea_ice_ts,order=c(0,1,0),seasonal=list(order=c(0,1,0), period=12))
res.m1 = residuals(m1_sea_ice);
plot(res.m1,xlab='Time',ylab='Residuals',main="Time series plot of the residuals")

```


# Note small error at line 241 from original TSA_Final_all,  res.m5 = residuals(m4.sea_ice);


## ---------------------------

##  https://www.methodology.psu.edu/resources/aic-vs-bic/

## Lowest AIC value is m7.sea_ice, SARIMA(0,1,2)x(0,1,1)~12~ model, 
## and                 m5.sea.ice  SARIMA(0,1,1)x(0,1,1)~12~ model

```{r}
##  Sort score special from Module 6 line 175 solutions and module 7 specials

sort.score <- function(x, score = c("bic", "aic")){
  if (score == "aic"){
    x[with(x, order(AIC)),]
  } else if (score == "bic") {
    x[with(x, order(BIC)),]
  } else {
    warning('score = "x" only accepts valid arguments ("aic","bic")')
  }
}

sort.score(AIC(m1_sea_ice, m2.sea_ice, m3.sea_ice,m4.sea_ice, m5.sea_ice, m6.sea_ice,m7.sea_ice, m8.sea_ice), score = "aic")
```

```{r}
library(forecast)
sort.score(BIC(m1_sea_ice, m2.sea_ice, m3.sea_ice,m4.sea_ice, m5.sea_ice, m6.sea_ice,m7.sea_ice, m8.sea_ice), score = "bic")
```



## While AIC shows m7 to be a good model , BIC shows that m5 would most likely be a better model 
## in that BIC is an estimate of a function of the posterior probability of a model being true

```{r}
eacf(res.m5)  ##  Looking at AR/MA for both m5 and m7
```

```{r}
eacf(res.m7)
```

```{r}
McLeod.Li.test(y=res.m5,main="McLeod-Li Test ")
```

```{r}
McLeod.Li.test(y=res.m7,main="McLeod-Li Test ")
```

```{r}

## From module 7 Specials 

residual.analysis <- function(model, std = TRUE,start = 2, class = c("ARIMA","GARCH","ARMA-GARCH")[1]){
  # If you have an output from arima() function use class = "ARIMA"
  # If you have an output from garch() function use class = "GARCH"
  # If you have an output from ugarchfit() function use class = "ARMA-GARCH"
  library(TSA)
  library(FitAR)
  if (class == "ARIMA"){
    if (std == TRUE){
      res.model = rstandard(model)
    }else{
      res.model = residuals(model)
    }
  }else if (class == "GARCH"){
    res.model = model$residuals[start:model$n.used]
  }else if (class == "ARMA-GARCH"){
      res.model = model@fit$residuals
  }else {
    stop("The argument 'class' must be either 'ARIMA' or 'GARCH' ")
  }
  par(mfrow=c(3,2))
  plot(res.model,type='o',ylab='Standardised residuals', main="Time series plot of standardised residuals")
  abline(h=0)
  hist(res.model,main="Histogram of standardised residuals")
  acf(res.model,main="ACF of standardised residuals")
  pacf(res.model,main="PACF of standardised residuals")
  qqnorm(res.model,main="QQ plot of standardised residuals")
  qqline(res.model, col = 2)
  print(shapiro.test(res.model))
  k=0
  LBQPlot(res.model, lag.max = 30, StartLag = k + 1, k = 0, SquaredQ = FALSE)
}

residual.analysis(model = m5.sea_ice)

```

```{r}
residual.analysis(model = m7.sea_ice)
```

##  Ljung is slightly better for m7

##  End here  --------------

##  From here on down is ramblings  ------------------------------------------


## Fit of SARIMA(0,1,1)x(0,1,1)~12~ model   <--  m1
## Fit of SARIMA(0,1,0)x(0,1,0)~12~ model   <--  m2
## Fit of SARIMA(2,1,0)x(0,1,1)~12~ model   <--  m3


```{r}
McLeod.Li.test(y=res.m5,main="McLeod-Li Test ")
```

```{r}
McLeod.Li.test(y=res.m7,main="McLeod-Li Test ")
```


```{r}

# ARIMA(0,1,3)
model_013_css = arima(sea_ice_ts ,order=c(1,1,3),method='CSS')
coeftest(model_013_css)

model_013_ml = arima(sea_ice_ts,order=c(0,1,3),method='ML')
coeftest(model_013_ml) 



#earnings.transform = BoxCox.ar(sea_ice_ts) # The default method for fitting is MLE here
# If do not get a smooth curve with MLE, you can change the method to least squares or Method of Moments
SeaIce.transform = BoxCox.ar(sea_ice_ts, method = 'ols')#yule-walker 
# earnings.transform = BoxCox.ar(earnings, method = "ols")
# , "burg", "ols", "mle", "yw"

SeaIce.transform$ci
SeaIce.log = log(sea_ice_ts) # 0 is in the interval or you can go for mid point of the interval as well

diff.earnings.log = diff(SeaIce.log,differences = 1)
plot(diff.earnings.log,type='o',ylab='Quarterly earnings ')

```


```{r}

# TS Handy: Handy Tools for Univariate Time Series Analysis
# Author(s): Yong Kai, Wong
# Version Number: 0.1
# Objective: To automate repetitive processes in time series analysis

## Install packages and load libraries
pkgs <- c('TSA', 'fUnitRoots', 'forecast', 'lmtest')
invisible(lapply(pkgs, require, character.only = T))

myCandidate <- function(timeSeries, orderList,
                        methodType = c("CSS-ML", "ML", "CSS")[1],
                        fixedList = NULL, includeConstant = c(TRUE, FALSE)[1]){
  
  # timeSeries = the time series (a ts object)
  # orderList = a list object of c(p, d, q)
  # methodType = estimation method; default = "CSS-ML"
  # fixedList = a list object of free/fixed coefficient
  # includeConstant = if true, an intercept term is included in the model
  
  myCandidateEst <- list()
  
  for(i in 1:length(orderList)){
    order <- sapply(orderList,function(x) unlist(x))[,i]
    myCandidateEst[[i]] <- Arima(y = timeSeries, order = order, method = methodType,
                                 include.constant = includeConstant)
  }
  return(myCandidateEst)
  
}

```


```{r}
modelList <- list(c(2,0,1), c(0,0,2), c(0,0,0), c(0,0,1),c(1,0,0), c(1,0,1),c(0,2,0), c(2,0,1),c(2,0,0), c(3,0,0))
modelEstimation <- myCandidate(res.m7, orderList = modelList, methodType = "ML")
# 
modelEstimation$IC
#  Not useable for seasonal
```







